\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-10-04 gwnnDraft02.tex
nitpicks of section 6

with $\epsilon$ being calculable from the scalar potential. Assuming that there is a negligible amount of energy being dumped back into the SM, which would cause significant reheating of $\rho_{\gamma}$,  the latent heat, $\epsilon$ should correspond to the energy density of the hidden sector going through the PT. The parameter $g_{*}$ in the denominator of Eq.~(\ref{eq:alpha}) is the number of relativistic degrees of freedom at the time of the phase transition, with contributions from species in both the visible and dark sectors. It has weak temperature dependence in a single sector, but when dealing with multiple hidden sectors, $g_{*}$ gains contributions from all $N$ sector's relativistic degrees of freedom, weighted by their respective energy densities

@@

Due to the lack of a general analytic QCD potential, it is not possible to use Eq.~(\ref{eqn:BH}) to calculate $\beta/H$. There are dimensional arguments~\cite{Hogan:1984hx,Hogan:1986qda} that predict $\beta/H \sim 4 \textrm{Log}(M_{p}/ \Lambda_{QCD,i})$, although these arguments make specific assumptions about the potential. In more recent work, some authors~\citep{Helmboldt:2019pan,Bai:2018dxf} have attempted to estimate it using first-order chiral effective theories and the Polyakov-Nambu-Jona-Lasinio (PNJL) models. These studies claim a large range of values and thus a consensus has not yet been reached.  Under these circumstances, we take the optimistic case in which $\beta/H$ is $\mathcal{O}(10)$.

@@

The exotic sectors have essentially massless degrees of freedom pre-phase transition and pions with negligible masses post-phase transition. Other composite particles, such as baryons, do gain a mass of the order of $\Lambda_{\rm QCD}$; this is, however, still much smaller than the order of $\rho_R$ leading to small $\alpha_{\infty}$ according to Eq.~(\ref{eqn:critPTstrength}).\footnote {It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors, their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$.} Heavy standard sectors that undergo SFOPT for QCD feature no baryons due to all quarks being above their respective QCD scales. They do, however, feature glueballs that obtain a mass of the order of the SFOPT and, just as in the case for exotic sectors above, feature small $\alpha_{\infty}$. Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Explicitly, all exotic sectors (and heavy standard sectors) have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). Each component of the spectral energy density in Eq.~(\ref{eqn:SpectralEng}) is proportional to a power of their respective efficiency factors $\kappa$. In this case, the efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$.

@@

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The Redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave spectral energy density (solid curves) for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded dashed curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity as described in Section~\ref{sec:detection}. The ones on the right are space-based interferometers: Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red). The ones on the left are for the pulsar timing array SKA~\cite{Janssen:2014dka} for exposure time of 5-year (purple), 10-year (orange), and 20-year (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, more detail in the construction of these curves is given in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sectors contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

@@

When using Eq.~(\ref{eqn::SNR}) to construct the PSI curves for SKA, the upper integration bound is inversely proportional to pulsar's timing residual, $f_{min} = 1/\delta t$.  In this work, it is assumed that $\delta t = 14$ days, but this may underestimate the capabilities of SKA as well as the cadences of the pulsar populations. If the timing residuals are lowered the maximum frequency reach of SKA increases, and the corresponding PLI curves in Figs.~\ref{fig::Nnatural} $\&$ \ref{fig:Haa} are shifted to the right, possibly giving the PTAs sensitivity to some of the scenarios considered here.

>>>>

with $\epsilon$ being calculable from the scalar potential. Assuming that there is a negligible amount of energy being dumped back into the SM, which would cause significant reheating of $\rho_{\gamma}$,  the latent heat $\epsilon$ should correspond to the energy density of the hidden sector going through the PT. The parameter $g_{*}$ in the denominator of Eq.~(\ref{eq:alpha}) is the number of relativistic degrees of freedom at the time of the phase transition, with contributions from species in both the visible and dark sectors. It has weak temperature dependence in a single sector, but when dealing with multiple hidden sectors, $g_{*}$ gains contributions from all $N$ sector's relativistic degrees of freedom, weighted by their respective energy densities

@@

Due to the lack of a general analytic QCD potential, it is not possible to use Eq.~(\ref{eqn:BH}) to calculate $\beta/H$. There are dimensional arguments~\cite{Hogan:1984hx,Hogan:1986qda} that predict $\beta/H \sim 4 \textrm{Log}(M_{p}/ \Lambda_{QCD,i})$, although these arguments make specific assumptions about the potential. In more recent work, some authors~\citep{Helmboldt:2019pan,Bai:2018dxf} have attempted to estimate it using first-order chiral effective theories and the Polyakov-Nambu-Jona-Lasinio (PNJL) models. These studies claim a large range of values with no consensus reached on the precise order of the scaled inverse timescale.  Under these circumstances, we take the optimistic case in which $\beta/H$ is $\mathcal{O}(10)$.

@@

The exotic sectors have essentially massless degrees of freedom pre-phase transition and pions with negligible masses post-phase transition. Other composite particles, such as baryons, do gain a mass of the order of $\Lambda_{ex}$; this is, however, still much smaller than the order of $\rho_R$ leading to small $\alpha_{\infty}$ according to Eq.~(\ref{eqn:critPTstrength}).\footnote {It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors, their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$.} Heavy standard sectors that undergo SFOPT for QCD feature no baryons due to all quarks being above their respective QCD scales. They do, however, feature glueballs that obtain a mass of the order of the SFOPT and, just as in the case for exotic sectors above, feature small $\alpha_{\infty}$. Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Explicitly, all exotic sectors (and heavy standard sectors) have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). Each component of the spectral energy density in Eq.~(\ref{eqn:SpectralEng}) is proportional to a power of their respective efficiency factors $\kappa$. In this case, the efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$.

@@

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave spectral energy density (solid curves) for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded dashed curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity as described in Section~\ref{sec:detection}. The ones on the right are space-based interferometers: Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red). The ones on the left are for the pulsar timing array SKA~\cite{Janssen:2014dka} for exposure time of 5-year (purple), 10-year (orange), and 20-year (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, more detail in the construction of these curves is given in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sectors contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

@@

When using Eq.~(\ref{eqn::SNR}) to construct the PLI curves for SKA, the upper integration bound is inversely proportional to pulsar's timing residual, $f_{min} = 1/\delta t$.  In this work, it is assumed that $\delta t = 14$ days, but this may underestimate the capabilities of SKA as well as the cadences of the pulsar populations. If the timing residuals are lowered the maximum frequency reach of SKA increases, and the corresponding PLI curves in Figs.~\ref{fig::Nnatural} $\&$ \ref{fig:Haa} are shifted to the right, possibly giving the PTAs sensitivity to some of the scenarios considered here.
